UMAP reveals cryptic population structure and phenotype heterogeneity in large genomic cohorts

### 2019-09-07 ms/main.tex
add e.g.

Methods such as UMAP and t-SNE focus on preserving local distances to reveal fine-scale structure in populations, and in the process may preserve aspects of global structure as well. In contrast, PCA preserves long range distances but hides finer-scale details. Hierarchical clustering of networks has also successfully detected fine-scale population structure using identity-by-descent similarity by attempting to preserve relations between global networks and smaller local ones\cite{han2017clustering}. We speculate that the addition of weak constraints favouring the preservation of longer distances in UMAP-like approaches has the potential of preserving the desirable local properties while encouraging more intuitive positioning of clusters on a global scale.

>>>>

Methods such as UMAP and t-SNE focus on preserving local distances to reveal fine-scale structure in populations, and in the process may preserve aspects of global structure as well. In contrast, PCA preserves long range distances but hides finer-scale details. Hierarchical clustering of networks has also successfully detected fine-scale population structure using identity-by-descent similarity by attempting to preserve relations between global networks and smaller local ones (e.g., \cite{han2017clustering}). We speculate that the addition of weak constraints favouring the preservation of longer distances in UMAP-like approaches has the potential of preserving the desirable local properties while encouraging more intuitive positioning of clusters on a global scale.
